\bf\mftitle

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial_platform.tex
Documentation: Tutorials: changing titles.
platform tut OK

\def\mftitle{Dealing with jInfer and NetBeans Platform}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

@@

\subsection*{Overview}
\begin{enumerate}
  \item Module visibility
  \item Error handling
  \item Interruptions
  \item Dialogs
  \item Configuration - options, preferences
  \item Rule display
  \item Console output, logging
  \item RunningProject class
  \item Module selection
\end{enumerate}
\subsection*{Module visibility}
\par 
  Each NBP module allows to set some spackages as public, which means their content
  will be available to other modules. That means, if module \textit{A} declares a
  dependency on module \textit{B}, it will be able to use only classes from
  those packages that were set as public in module \textit{B}.

@@

Example:
\begin{verbatim}
// This is a portion of example project.xml file.
<?xml version="1.0" encoding="UTF-8"?>
<project xmlns="\url{http://www.netbeans.org/ns/project/1">}
    <type>org.netbeans.modules.apisupport.project</type>
    <configuration>
        <data xmlns="\url{http://www.netbeans.org/ns/nb-module-project/3">}
            ....
            <public-packages>
                <package>some.example.package.name</package>
            </public-packages>
        </data>
    </configuration>
</project>
\end{verbatim}
\par 
  Another way to set public packages is through GUI:
  \begin{enumerate}
    \item Right-click the module, select \textit{Properties}.
    \item In \textit{API Versioning}, there is a \textit{Public Packages} section.
    \item Check all packages that should be public.
  \end{enumerate}

@@

Example:
\begin{verbatim}
for (forever) {
    if (Thread.interrupted()) {
        throw new InterruptedException();
    }
    doStuff();
}
\end{verbatim}
\subsection*{Dialogs}
\par 
  Creation of open/save dialogs or message windows is done using standard NBP API.
  For dialogs you can use the standard Java \texttt{JFileChooser},
  but NBP \texttt{FileChooserBuilder}
  is more convenient. This class creates dialog which remembers last-used
  directory according to a key passed into its constructor.

Example:
\begin{verbatim}
// The default dir to use if no value is stored
File home = new File(System.getProperty("user.home"));
// Now build a file chooser and invoke the dialog in one line of code
File toAdd = new FileChooserBuilder(CallingClass.class)
                   .setTitle("Some dialog title")
                   .setDefaultWorkingDirectory(home).showOpenDialog();
// Result will be null if the user clicked cancel or closed the dialog w/o OK
if (toAdd != null) {
    //do something
}
\end{verbatim}
\par 
  For message windows the \texttt{DialogDisplayer} NBP API
  class is used. To determine which type of message will be shown
  (Confirmation dialog, message dialog...),
  \texttt{NotifyDescriptor}
  is used. Following is a small example, which creates standard message
  notification. For more information please look at
  NBP Dialogs API FAQ.

Example:
\begin{verbatim}
// Creates standard information message with text "Hello world"
NotifyDescriptor nd = new NotifyDescriptor
                            .Message("Hello world",
                               NotifyDescriptor.INFORMATION_MESSAGE);
DialogDisplayer.getDefault().notify(nd);
\end{verbatim}
\subsection*{Configuration - options, preferences}
\par 
  Option panels located in \textit{Tools > Options} menu are a standard part of
  NBP API. We created a \textit{jInfer} option sub-category to place panels with
  configuration valid across all jInfer projects. Description of the few step
  to create your own option panel follows. For more information, please visit
  NetBeans Options tutorial.

@@

Example:
\begin{verbatim}
public void store() {
  // Saves the state of checkBox into Preferences for ExampleClass
  // class in property with name "PropertyName"
  NbPreferences.forModule(ExampleClass.class)
      .putBoolean("PropertyName", someCheckBox.isSelected());
}
public void load() {
  // Get from ExampleClass class preferences the property with name
  // "PropertyName". Second parameter of method getBoolean is used
  // as default if no property with defined name is saved in preferences.
  someCheckBox.setSelected(NbPreferences.forModule(ExampleClass.class)
                               .getBoolean("PropertyName", true));
}
\end{verbatim}
\par 
  Second type of preferences used in jInfer is Project properties. Each jInfer
  project has its properties panel accessible in its content menu and
  these properties are applied for each project separately. To implement this
  kind of preferences, it is necessary to implement
  \texttt{PropertiesPanelProvider}
  interface and extend the
  \texttt{AbstractPropertiesPanel}
  class. Each Project properties window has category tree, where each category
  represents a separate panel with properties. This category is declared in the
  provider interface, properties panel itself is defined in class extending
  \texttt{AbstractPropertiesPanel}.

\paragraph*{\texttt{PropertiesPanelProvider}}
\par 
  As was mentioned above, \texttt{PropertiesPanelProvider} defines
  category in Project properties windows. In the example below, we'll try to
  explain how to create a simple provider and what each method is responsible for.

Example:
\begin{verbatim}
public class ExampleProvider implemenets PropertiesPanelProvider {
  // Programmatic name of this provider.
  public String getName() {
    return "ExampleProvider";
  }
  // User-friendly name - this name will be displayed in category tree.
  public String getDisplayName() {
    return "Example Category";
  }
  // Priority of category. Higher the number, higher
  // will the category be in the tree. If two categories have
  // same priorities, they are sorted according
  // to their names.
  public int getPriority() {
    return 0;
  }
  // Returns properties panel defined for this category.
  public AbstractPropertiesPanel getPanel(final Properties properties) {
    return new ExamplePropertiesPanel(properties);
  }

  // Defines parent category of this category. If null is returned, this is top
  // level category. In other cases id (programmatic name) of parent category
  // must be returned.
  public String getParent() {
    return null;
  }

  // Optional. For each category, list of virtual categories can be defined.
  // Instead of properties panel, VirtualCategoryPanel only informs of type and
  // installed number of modules of a certain type.
  public List getSubCategories() {
    return null;
  }
}
\end{verbatim}

\paragraph*{\texttt{AbstractPropertiesPanel}}
\par 
  This class represents the visual component of properties category. You can
  design it in whatever way you like. For proper functionality, a few steps must
  be followed.

@@

Example:
\begin{verbatim}
// This is only a part of the actual class...
public class ExamplePanel extends AbstractPropertiesPanel {

@@

  public abstract void load() {
    String propValue = properties.getProperty("exampleKey", "default value");
    someTextField.setText(propValue);
  }
}
\end{verbatim}

\subsection*{Rule display}

\par 
  \textit{Rule displayer} is a component used for visualization of rules in any
  step of inference process. If rule displayers bundled with jInfer are
  insufficient, you can implement a rule displayer on your own.
  You just need to follow these few easy steps.

@@

\par 
  \textbf{Important notice:} never forget to clone the rules you want
  to display. Rule displayer usually runs in its own thread, and accessing the
  rules that are being simplified at the same time might lead to weird results.
  You can use \texttt{CloneHelper} to clone the rules.

Example:
\begin{verbatim}
@ServiceProvider(service = RuleDisplayer.class)
public final class ExampleRuleDisplayer implements RuleDisplayer {

@@

  @Override
  public String getModuleDescription() {
    return getDisplayName();
  }
}
\end{verbatim}

Example of invocation with cloning:
\begin{verbatim}
RuleDisplayerHelper.showRulesAsync("Panel name",
    new CloneHelper().cloneGrammar(grammar), true);
\end{verbatim}

@@

Example:
\begin{verbatim}
// Creates new output window, where first parameter is a name and second
// is flag determining whether a new window should be created even if there is
//  already a window with the same name.
InputOutput ioResult = IOProvider.getDefault().getIO("Example", true);
ioResult.getOut().println("Hello world");

// After writing everything into output window, close the output.
ioResult.getOut().close();
\end{verbatim}

\par 
  Logging is done in jInfer using
  Log4j tool.
  If you are not familiar with Log4j, example below shows simple usage. All
  logged messages are saved in a standard log file placed in \textit{UserHome/.jinfer/}
  folder and printed into "jInfer" output window in the application. For each
  of these two places a log level can be set in jInfer options.

Example:
\begin{verbatim}
// Creates Logger instance for this particular class
private static final Logger LOG = Logger.getLogger(Example.class);

@@

\subsection*{\texttt{RunningProject} class}

@@

\begin{verbatim}
public interface Foo extends NamedModule {

  String bar();
}
\end{verbatim}

@@

Example of the first implementation. The second will be the same (just change "ones" to "twos").
\begin{verbatim}
@ServiceProvider(service = Foo.class)
public class Impl1 implements Foo {

@@

Example of panel load code:
\begin{verbatim}
@Override
public final void load() {
  combo.setModel(new DefaultComboBoxModel(
          ModuleSelectionHelper.lookupImpls(Foo.class).toArray()));

@@

Example of panel save code:
\begin{verbatim}
@Override
public void store() {
  properties.setProperty("selected.foo.implementation",
          ((NamedModule) combo.getSelectedItem()).getName());
}
\end{verbatim}

Example of combo box renderer setting:
\begin{verbatim}
combo.setRenderer(new ProjectPropsComboRenderer(combo.getRenderer()));
\end{verbatim}

>>>>

\def\mftitle{jInfer: Dealing with jInfer and NetBeans Platform}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

@@

\subsection*{Overview}
\begin{enumerate}
  \item Module visibility
  \item Error handling
  \item Interruptions
  \item Dialogs
  \item Configuration - options, preferences
  \item Rule display
  \item Console output, logging
  \item RunningProject class
  \item Module selection
\end{enumerate}
\subsection*{Module visibility}
\par 
  Each NBP module allows to set some packages as public, which means their content
  will be available to other modules. That means, if module \textit{A} declares a
  dependency on module \textit{B}, it will be able to use only classes from
  those packages that were set as public in module \textit{B}.

@@

\noindent Example:
\begin{verbatim}
// This is a portion of example project.xml file.
<?xml version="1.0" encoding="UTF-8"?>
<project xmlns="http://www.netbeans.org/ns/project/1">
    <type>org.netbeans.modules.apisupport.project</type>
    <configuration>
        <data xmlns="http://www.netbeans.org/ns/nb-module-project/3">
            ....
            <public-packages>
                <package>some.example.package.name</package>
            </public-packages>
        </data>
    </configuration>
</project>
\end{verbatim}
\par 
\noindent  Another way to set public packages is through GUI:
  \begin{enumerate}
    \item Right-click the module, select \textit{Properties}.
    \item In \textit{API Versioning}, there is a \textit{Public Packages} section.
    \item Check all packages that should be public.
  \end{enumerate}

@@

\noindent Example:
\begin{verbatim}
for (forever) {
    if (Thread.interrupted()) {
        throw new InterruptedException();
    }
    doStuff();
}
\end{verbatim}
\subsection*{Dialogs}
\par 
  Creation of open/save dialogs or message windows is done using standard NBP API.
  For dialogs you can use the standard Java \texttt{JFileChooser},
  but NBP \texttt{FileChooserBuilder}
  is more convenient. This class creates dialog which remembers last-used
  directory according to a key passed into its constructor.

\noindent Example:
\begin{verbatim}
// The default dir to use if no value is stored
File home = new File(System.getProperty("user.home"));
// Now build a file chooser and invoke the dialog in one line of code
File toAdd = new FileChooserBuilder(CallingClass.class)
                   .setTitle("Some dialog title")
                   .setDefaultWorkingDirectory(home).showOpenDialog();
// Result will be null if the user clicked cancel or closed the dialog w/o OK
if (toAdd != null) {
    //do something
}
\end{verbatim}
\par 
  For message windows the \texttt{DialogDisplayer} NBP API
  class is used. To determine which type of message will be shown
  (Confirmation dialog, message dialog...),
  \texttt{NotifyDescriptor}
  is used. Following is a small example, which creates standard message
  notification. For more information please look at
  NBP Dialogs API FAQ.

\noindent Example:
\begin{verbatim}
// Creates standard information message with text "Hello world"
NotifyDescriptor nd = new NotifyDescriptor
                            .Message("Hello world",
                               NotifyDescriptor.INFORMATION_MESSAGE);
DialogDisplayer.getDefault().notify(nd);
\end{verbatim}
\subsection*{Configuration - options, preferences}
\par 
  Option panels located in \textit{Tools > Options} menu are a standard part of
  NBP API. We created a \textit{jInfer} option sub-category to place panels with
  configuration valid across all jInfer projects. Description of the few step
  to create your own option panel follows. For more information, please visit
  NetBeans Options tutorial.

@@

\noindent Example:
\begin{verbatim}
public void store() {
  // Saves the state of checkBox into Preferences for ExampleClass
  // class in property with name "PropertyName"
  NbPreferences.forModule(ExampleClass.class)
      .putBoolean("PropertyName", someCheckBox.isSelected());
}

public void load() {
  // Get from ExampleClass class preferences the property with name
  // "PropertyName". Second parameter of method getBoolean is used
  // as default if no property with defined name is saved in preferences.
  someCheckBox.setSelected(NbPreferences.forModule(ExampleClass.class)
                               .getBoolean("PropertyName", true));
}
\end{verbatim}
\par 
  Second type of preferences used in jInfer is Project properties. Each jInfer
  project has its properties panel accessible in its content menu and
  these properties are applied for each project separately. To implement this
  kind of preferences, it is necessary to implement
  \texttt{Properties\.Panel\.Provider}
  interface and extend the
  \texttt{Abstract\.Properties\.Panel}
  class. Each Project properties window has category tree, where each category
  represents a separate panel with properties. This category is declared in the
  provider interface, properties panel itself is defined in class extending
  \texttt{AbstractPropertiesPanel}.

\subsubsection*{PropertiesPanelProvider}
\par 
  As was mentioned above, \texttt{PropertiesPanelProvider} defines
  category in Project properties windows. In the example below, we'll try to
  explain how to create a simple provider and what each method is responsible for.

\noindent Example:	
\begin{verbatim}
public class ExampleProvider implemenets PropertiesPanelProvider {
  // Programmatic name of this provider.
  public String getName() {
    return "ExampleProvider";
  }
  // User-friendly name - this name will be displayed in category tree.
  public String getDisplayName() {
    return "Example Category";
  }
  // Priority of category. Higher the number, higher
  // will the category be in the tree. If two categories have
  // same priorities, they are sorted according
  // to their names.
  public int getPriority() {
    return 0;
  }
  // Returns properties panel defined for this category.
  public AbstractPropertiesPanel getPanel(final Properties properties) {
    return new ExamplePropertiesPanel(properties);
  }
  // Defines parent category of this category. If null is returned, this is top
  // level category. In other cases id (programmatic name) of parent category
  // must be returned.
  public String getParent() {
    return null;
  }
  // Optional. For each category, list of virtual categories can be defined.
  // Instead of properties panel, VirtualCategoryPanel only informs of type and
  // installed number of modules of a certain type.
  public List getSubCategories() {
    return null;
  }
}
\end{verbatim}

\subsubsection*{AbstractPropertiesPanel}
\par 
  This class represents the visual component of properties category. You can
  design it in whatever way you like. For proper functionality, a few steps must
  be followed.

@@

\end{enumerate}
\noindent Example:
\begin{verbatim}
// This is only a part of the actual class...
public class ExamplePanel extends AbstractPropertiesPanel {

@@

  public abstract void load() {
    String propValue = properties.getProperty("exampleKey", "default value");
    someTextField.setText(propValue);
  }
}
\end{verbatim}
\newpage

\subsection*{Rule display}
  \textit{Rule displayer} is a component used for visualization of rules in any
  step of inference process. If rule displayers bundled with jInfer are
  insufficient, you can implement a rule displayer on your own.
  You just need to follow these few easy steps.

@@

\par \noindent
  \textbf{Important notice:} never forget to clone the rules you want
  to display. Rule displayer usually runs in its own thread, and accessing the
  rules that are being simplified at the same time might lead to weird results.
  You can use \texttt{CloneHelper} to clone the rules.

\noindent Example:
\begin{verbatim}
@ServiceProvider(service = RuleDisplayer.class)
public final class ExampleRuleDisplayer implements RuleDisplayer {

@@

  @Override
  public String getModuleDescription() {
    return getDisplayName();
  }
}
\end{verbatim}
Example of invocation with cloning:
\begin{verbatim}
RuleDisplayerHelper.showRulesAsync("Panel name",
    new CloneHelper().cloneGrammar(grammar), true);
\end{verbatim}

@@

\noindent Example:
\begin{verbatim}
// Creates new output window, where first parameter is a name and second
// is flag determining whether a new window should be created even if there is
//  already a window with the same name.
InputOutput ioResult = IOProvider.getDefault().getIO("Example", true);
ioResult.getOut().println("Hello world");

// After writing everything into output window, close the output.
ioResult.getOut().close();
\end{verbatim}
  Logging is done in jInfer using
  Log4j tool.
  If you are not familiar with Log4j, example below shows simple usage. All
  logged messages are saved in a standard log file placed in \textit{UserHome/.jinfer/}
  folder and printed into ``jInfer'' output window in the application. For each
  of these two places a log level can be set in jInfer options.

\noindent Example:
\begin{verbatim}
// Creates Logger instance for this particular class
private static final Logger LOG = Logger.getLogger(Example.class);

@@

\subsection*{RunningProject class}

@@

\begin{verbatim}
public interface Foo extends NamedModule {
  String bar();
}
\end{verbatim}

@@

\noindent Example of the first implementation. The second will be the same (just change ``ones'' to ``twos'').
\begin{verbatim}
@ServiceProvider(service = Foo.class)
public class Impl1 implements Foo {

@@

\noindent Example of panel load code:
\begin{verbatim}
@Override
public final void load() {
  combo.setModel(new DefaultComboBoxModel(
          ModuleSelectionHelper.lookupImpls(Foo.class).toArray()));

@@

\noindent Example of panel save code:
\begin{verbatim}
@Override
public void store() {
  properties.setProperty("selected.foo.implementation",
          ((NamedModule) combo.getSelectedItem()).getName());
}
\end{verbatim}

\noindent Example of combo box renderer setting:
\begin{verbatim}
combo.setRenderer(new ProjectPropsComboRenderer(combo.getRenderer()));
\end{verbatim}
